\propertitle

### 2019-04-23 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-20T23:54:37+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\usepackage{wrapfig}

@@

\section{Why polls?}
\label{sec:polls}

@@

\section{Maximum-entropy method}
\label{sec:maxent_remarks}

A necessary step of every plausible inference is the assignment of some
probabilistic premisses; as many as necessary to arrive at our desired
probabilistic conclusions given the evidence. The necessity of this step
was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
a century ago:\vspace{\topsep}
\begin{quote}\footnotesize
  Now the axioms of probability enable us to infer any
  probability-conclusion \emph{only} from probability-premisses. In other
  words, the calculus of probability does not enable us to infer any
  probability-value unless we have some probabilities or probability
  relations \emph{given}. Such data cannot be supplied by the
  mathematician. E.g. the rules of arithmetic and the axioms of the
  probability-calculus are utterly impotent to determine, on the supposed
  knowledge that the throw of a coin must yield either head or tail and
  cannot yield both, the probability that it will yield head or that it
  will yield tail. We must assume that the two co-exclusive and
  co-exhaustive possibilities are \emph{equally probable}, before we can
  estimate the probability of either as being a half of certitude.
  \sourceatright{\citep[Appendix on eduction, \sect~5,
    p.~182]{johnson1924}}
\end{quote}

These premisses may represent either actual knowledge or just
\setlength{\intextsep}{0.5ex}% with wrapfigure
\begin{wrapfigure}{l}{0.33\linewidth}% with wrapfigure
\centering\includegraphics[width=\linewidth]{statements_scheme1.pdf}%\\
%\caption{caption}\label{fig:comparison_a5}
\end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
some hypotheses whose consequences we want to explore. They can be assigned
at different depths of the logical analysis of our inference problem.

We can for example assign probabilistic premisses at a very deep analytical
level, as schematically illustrated on the left. They determine the
probabilities of statements at more superficial levels, including the
statements which we're ultimately interested in or which we can directly
observe.

\begin{wrapfigure}{r}{0.33\linewidth}% with wrapfigure
\centering\includegraphics[width=\linewidth]{statements_scheme2b.pdf}%\\
%\caption{caption}\label{fig:comparison_a5}
\end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
Or we can assign probabilistic premisses at a level of superficial or
intermediate depth, as schematically illustrated on the right. In this case
they determine the probabilities of statements at more superficial levels, and
partly constrain the probabilities of statements at deeper levels.

The choice of the analytical depth at which to assign the premisses depends
on many factors: on the difficulty of assessing our state of uncertainty at
that depth, on the cost of the ensuing calculations, and especially on the
general purpose of our inference. Assignments of probabilistic premisses at
different levels can still turn out to be completely equivalent for the
probabilities of the statements of interest.

Here's an example. Suppose that $n$ neurons have been recorded in a
sequence of time bins. We're given the time averages of the numbers of
active neurons and of neuron pairs in the sequence. We don't know the exact
sequence of activities or the number $T$ of time bins, although we know
that the latter is large. We want to forecast how many of the $n$ neurons
will be active in a \emph{new} time bin.

In a deep approach to this inference problem we assign the probability for
the possible value of $T$, and the probabilities for several hypotheses
about the sequence of activities in $T+1$ or more time bins. One hypothesis
could be, for instance, that the $n$ neurons are a sample of a larger
population of $N$ neurons, and a second hypothesis could be that for this
population there are some sufficient statistics, out of biological reasons.
We could of course deepen this approach further, with no end, down to
assigning the probabilities for the constitution and motion of the
molecules of the neurons and their environment, the probabilities for the
position and response of the recording electrodes, and so on. From all
these probabilities and the data we could finally calculate the probability
for the activity in the new time bin.

At the other end, the shallowest approach to this inference problem is to
directly assign the probabilities for the number of active neurons in the
new time bin, somehow using the data we were given. (It must be remarked
that such assignment partly constrains, in a backward fashion, all
assignments we could do at a deeper level.)

The maximum-entropy method was proposed by Jaynes
\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
premisses at a more direct level, as in the approach just described. This
point is made especially clear in \textcite{jaynes1994_r2003},
\sect~11.1:\vspace{0.5\topsep}
\begin{quote}\footnotesize
  If we knew the numbers $N$, $L$, $W$, \textins{some quantities relevant
    to our inference} then this could be solved by direct application of
  Bayes' theorem; without that information, we could still introduce the
  unknowns $N$, $L$, $W$ as nuisance parameters and use Bayes' theorem,
  eliminating them at the end. \textelp{} However, the Bayesian solution
  would not really address our problem; it only transfers it to the problem
  of assigning priors to $N$, $L$, $W$, leaving us back in essentially the
  same situation; how do we assign informative probabilities?
\end{quote}
The idea behind this method is to translate \emph{all} the data we're
given into constraints -- such as expectations -- for our probability
assignment, resolving the leftover freedom by choosing the constrained
probability distribution having highest Shannon entropy, as a measure of
its uncertainty or broadness. It can be shown that the use of this
method is approximately equivalent, under specific conditions, to a
probability assignment at a deeper level based on an assumption of
exchangeability together with a multinomial probability distribution for
long-run frequencies, or with an assumption about sufficient statistics
\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2017,portamana2009}.

Sometimes this method is applied with only \emph{part} of the available
data used as constraint; such an altered approach can be a heuristic way of
testing for sufficient statistics.

Owing to its directness, the maximum-entropy method is straightforward
to use -- and to misuse. In our previous example, for instance, it would
give silly results if the number of time bins were not much larger than the
possible activity values (it has thus been variously misused in the
neuroscientific literature \mynote{refs, especially to Tkacik}).

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-23T09:55:00+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

%\usepackage{wrapfig}

@@

\section{Polling neurons}
\label{sec:polls}

@@
